The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-09-30 paper.tex
Changed SonAngle abbrev to Son$\measuredangle$ and miscellaneous edits

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[table]{xcolor}    %for shading*
\usepackage{fullpage}
\usepackage{natbib}
\usepackage{mathpazo}
\usepackage{tipa}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{linguex}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage{slashbox}	
\usepackage{pifont}    %for pointing hand
\usepackage{arydshln}  %for dashed lines
\usepackage{rotating}  %for angled text
\usepackage{wasysym}   %for frowny face

@@

In this paper, I adopt (with, later, minor modifications) the following standard scale:

@@

I formalise the idea of faithfulness cost by defining a family of \textsc{Ident} constraints that penalise outputs that incur faithfulness costs of greater than a certain $n$, following \citep{flemming.2008}.

\ex. \textsc{Ident(SonAngle)}$<$n: Assign a violation mark if the consonants in two strings C\textsubscript{1}C\textsubscript{2} and C\textsubscript{1}VC\textsubscript{2} 
stand in correspondence, and the sonority angle between C\textsubscript{1}C\textsubscript{2} and C\textsubscript{1}V is greater than $n$.

This family of faithfulness constraints forms a stringency hierarchy. Any cluster that would violate \textsc{Ident(SonAngle)}$<$n, would also violate {\sc Ident(SonAngle)}$<$m, where m$>$n.

@@

Similarly, between the clusters that fall in sonority by two steps -- RF and NT -- we expect
NT to be less likely to undergo epenthesis, since N is less sonorous than R. We expect NT and RT to be the clusters most resistant to epenthesis out of all the clusters. This will be crucial 
to our analysis of Irish sonority-driven epenthesis.

@@

There are two possible ways to rank the family of {\sc Ident(SonAngle)}$<$n constraints.
One is to fix them in a universal ranking \citep{prince.smolensky.1993}, with least stringent the highest-ranked:

\ex. Universal ranking: \\
     {\sc Ident(SonAngle)}$<$2.0 $\gg$ {\sc Ident(SonAngle)}$<$1.5 $\gg$ ...

@@

\ex. Possible free ranking:
     {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Ident(SonAngle)}$<$2.0 $\gg$ ...

Depending on what constraints intervene, this may result in conflation of certain areas of the hierarchy, so that any distinctions between {\sc Sonority Angles} between 1.5 and 2.0 may be conflated.

@@

In this paper, for ease of analysis, I will assume that all {\sc Ident(SonAngle)}$<$n constraints are ranked low by default, but that individual {\sc Ident(SonAngle)}$<$n constraints can be promoted to dominate a markedness constraint $\mathcal{M}$ against clusters or another faithfulness constraint $\mathcal{F}$, blocking epenthesis for all clusters with a {\sc Sonority Angle} greater than $n$, with the result that either the cluster remains intact, or a competing solution to epenthesis emerges.

An example of {\sc Ident(SonAngle)}$<$n blocking epenthesis is given in the sample tableaux below.

\ex. Ident(SonAngle)$<$1.5 $\gg$ {\sc *ComplexCoda} $\gg$ {\sc Dep}
     \a. Allows epenthesis in rising and level clusters such as stop-fricative TF:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/TF/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline a. & & TF & & $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & T\textipa{@}F & 0.59 & & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. Prevents epenthesis in falling clusters such as FT:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline c. & \ding{43} & FT & &\cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} \\
\hline d. & & F\textipa{@}T & 2.11 $\ast$! &\cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

I will illustrate the use of {\sc Ident(SonAngle)}$<$n to block epenthesis in favour of other cluster resolution strategies with an example from the loanword phonology of the Dhaka dialect of Bangla, discussed and analysed in \citep{karim.2011}. Standard Bangla keeps coda clusters in words borrowed from English, but these are simplified in the Dhaka dialect in a variety of ways, including both epenthesis and deletion of one of the consonants in the cluster.

@@

However, if we retain this ranking, we incorrectly expect RT clusters to be broken up by epenthesis rather than delete the [r]. This can be resolved by ranking {\sc Ident(SonAngle)}$<$2.3, say, above {\sc Max}(r). This then blocks epenthesis in RT clusters, which have a {\sc Sonority Angle} of 2.35:

\ex. {\sc *ComplexCoda}, {\sc Ident(SonAngle)}$<$2.3  $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{park}/} & {\sc *ComplexCoda} & {\sc Ident(SonAngle)}$<$2.3 & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{park} & $\ast$! & & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline b. & & \textipa{parak} & & $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline c. & \ding{43} & \textipa{pak} & & & \cellcolor{lightgray}$\ast$ & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

While other analyses could possibly account for this data (see one such in \citep{karim.2011}), we see that simply ranking {\sc Ident(SonAngle)}$<$n $\gg$ $\mathcal{F}$ allows us to account for it quite straightforwardly.

@@

Thus far, we have not made any particular use of free ranking. While ranking {\sc Ident(SonAngle)}$<$n $\gg$ $\mathcal{M}$ does conflate all clusters with a {\sc Sonority Angle} above $n$, as well as all clusters with a {\sc Sonority Angle} below $n$, preserving only the distinction between {\sc Sonority Angles} greater and less than $n$, this could be accomplished using a universal ranking of the {\sc Ident(SonAngle)}$<$n constraints as well, as illustrated in the following example:

\ex. The {\sc Ident(SonAngle)}$<$n with the smallest $n$ that outranks {\sc *ComplexCoda} determines the cut-off point.

@@

Suppose a triconsonantal cluster /CCC/ has to be broken up either as [C\textipa{@}CC] or [CC\textipa{@}C], and further that [C\textipa{@}CC] violates some additional markedness constraint $\mathcal{M}$. Which {\sc Ident(SonAngle)}$<$n constraint outranks $\mathcal{M}$ will then determine the positioning of the epenthetic vowel.

\ex. {\sc Ident(SonAngle)}$<$n $\gg$ $\mathcal{M}$
     \a. Scenario 1: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
         Since neither violates {\sc Ident(SonAngle)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(SonAngle)}$<$n & $\mathcal{M}$ \\[0.5ex]
\hline \hline a. & \ding{43} & C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $<$n & \\
\hline b. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $<$n & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
      \b. Scenario 2: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$>$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
          C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} harmonically bound by C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} candidate.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(SonAngle)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline c. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}  & $<$n & \cellcolor{lightgray} \\
\hline d. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Scenario 3: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$<$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$>$n \\
         {\sc Ident(SonAngle)}$<$n determines output, overruling $\mathcal{M}$.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(SonAngle)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline e. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$!  & \cellcolor{lightgray} \\
\hline f. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3}& $>$n & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \d. Scenario 4: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2}-C\textsubscript{3}) $>$ n \\
         Since both candidates violate {\sc Ident(SonAngle)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(SonAngle)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline g. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$  & \\
\hline h. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$ & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Because we have assumed that all other {\sc Ident(SonAngle)}$<$n constraints rank low, certainly below $\mathcal{M}$, the degree to which candidates (g) and (h) violate {\sc Ident(SonAngle)}$<$n is irrelevant --- the distinction between clusters with {\sc Sonority Angles} above $n$ has been conflated. Even if {\sc Sonority Angle}(C\textsubscript{2},C\textsubscript{3}) exceeds {\sc Sonority Angle}(C\textsubscript{1},C\textsubscript{2}), the outcome is determined by the markedness constraint $\mathcal{M}$.

@@

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(SonAngle)}$<$1.9 & {\sc Ident(SonAngle)}$<$1.5 &$ \mathcal{M}$  \\[0.5ex]
\hline \hline a. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} &  2.0 $\ast$!  & \cellcolor{lightgray} 2.0 $\ast$ & \cellcolor{lightgray} \\
\hline b.  & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} &  1.7  & \cellcolor{lightgray} 1.7 $\ast$ & \cellcolor{lightgray} $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Under a universal ranking of {\sc Ident(SonAngle)}$<$n constraints, [C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}] would win in this scenario. In an equivalent scenario under free-ranking, where {\sc Ident(SonAngle)}$<$1.5 is the least stringent constraint to outrank $\mathcal{M}$, which in turn outranks all more stringent {\sc Ident(SonAngle)}$<$n constraints, the [C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}] candidate would win instead. There would be conflation of all clusters with {\sc Sonority Angles} below $n$, but distinctions between clusters with {\sc Sonority Angles} would still be relevant.

The example of Chaha will show that free-ranking is necessary, as conflation is observed on both sides of $n$, where {\sc Ident(SonAngle)}$<$n is the most stringent constraint to outrank $\mathcal{M}$.

@@

As with {\sc Sonority Angle}, this gives rise to a family of constraints {\sc Ident(SonRise)}$<$n, defined similarly to {\sc Ident(SonAngle)}$<$n.

@@

The fact that this process of epenthesis does not target obstruent-initial clusters is not problematic under the sonority contour faithfulness approach,
as the {\sc Ident(SonAngle)} constraints limit epenthesis, rather than trigger it.  By choosing a more limited markedness constraint such as {\sc *Son-C},
defined below, we can have epenthesis act only on sonorant-initial clusters.

@@

Under a sonority contour faithfulness analysis, epenthesis is triggered by a markedness constraint on the structural conditions of the cluster, not by the markedness of the underlying sonority contour profile itself. Undominated {\sc *ComplexOnset} accounts for the non-existence of initial clusters. {\sc *Coda} is ranked low, and therefore medial clusters are left intact.{\sc *ComplexCoda}, dominated by an {\sc Ident(SonAngle)} constraint, explains why some coda clusters are broken up by epenthesis while others are not.

@@

\ex. {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc *ComplexCoda}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{k1tf}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc *ComplexCoda} \\[0.5ex]
\hline \hline a. & & \textipa{k1tf} & & $\ast$! \\
\hline b. & \ding{43} & \textipa{k1t1f} & 0.59 & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{k1rm}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc *ComplexCoda} \\[0.5ex]
\hline \hline c. &  \ding{43} & \textipa{k1rm} & & $\ast$ \\
\hline d. & & \textipa{k1r1m} & 1.89 $\ast$! & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc Ident(SonAngle)}$<$1.9 $\gg$ {\sc *ComplexCoda}

>>>>

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[table]{xcolor}    %for shading*
\usepackage{fullpage}
\usepackage{natbib}
\usepackage{mathpazo}
\usepackage{tipa}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{linguex}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{tkz-euclide}
\usetkzobj{all}
\usepackage{slashbox}	
\usepackage{pifont}    %for pointing hand
\usepackage{arydshln}  %for dashed lines
\usepackage{rotating}  %for angled text
\usepackage{wasysym}   %for frowny face
\usepackage{amssymb} % for angle symbol

@@

In this paper, I adopt (with, later, minor modifications) the following standard scale \citep[(10)]{flemming.2008}:

@@

I formalise the idea of faithfulness cost by defining a family of \textsc{Ident} constraints that penalise outputs that incur faithfulness costs of greater than a certain $n$, following \cite{flemming.2008}.

\ex. \textsc{Ident(Son$\measuredangle$)}$<$n: Assign a violation mark if the consonants in two strings C\textsubscript{1}C\textsubscript{2} and C\textsubscript{1}VC\textsubscript{2} 
stand in correspondence, and the {\sc Sonority Angle} between C\textsubscript{1}C\textsubscript{2} and C\textsubscript{1}V is greater than $n$.

This family of faithfulness constraints forms a stringency hierarchy. Any cluster that would violate \textsc{Ident(Son$\measuredangle$)}$<$n, would also violate {\sc Ident(Son$\measuredangle$)}$<$m, where m$>$n.

@@

Similarly, between the clusters that fall in sonority by two steps --- RF and NT --- we expect
NT to be less likely to undergo epenthesis, since N is less sonorous than R.

Among falling clusters that share the same C\textsubscript{2}, there is a tension between wanting a small sonority distance and wanting C\textsubscript{1} to have as high a sonority as possible. RT and NT resolve this in different ways, ending up with the same {\sc Sonority Angle}, which is the largest out of the all the clusters. We thus expect NT and RT to be the clusters most resistant to epenthesis out of all the clusters. This will be crucial to our analysis of Irish sonority-driven epenthesis.

@@

There are two possible ways to rank the family of {\sc Ident(Son$\measuredangle$)}$<$n constraints.
One is to fix them in a universal ranking \citep{prince.smolensky.1993}, with least stringent the highest-ranked:

\ex. Universal ranking: \\
     {\sc Ident(Son$\measuredangle$)}$<$2.0 $\gg$ {\sc Ident(Son$\measuredangle$)}$<$1.5 $\gg$ ...

@@

\ex. Possible free ranking:
     {\sc Ident(Son$\measuredangle$)}$<$1.5 $\gg$ {\sc Ident(Son$\measuredangle$)}$<$2.0 $\gg$ ...

Depending on what constraints intervene, this may result in conflation of certain areas of the hierarchy, so that any distinctions between {\sc Sonority Angles} between 1.5 and 2.0 are made irrelevant.

@@

In this paper, for ease of analysis, I will assume that all {\sc Ident(Son$\measuredangle$)}$<$n constraints are ranked low by default, but that individual {\sc Ident(Son$\measuredangle$)}$<$n constraints can be promoted to dominate a markedness constraint $\mathcal{M}$ against clusters or another faithfulness constraint $\mathcal{F}$, blocking epenthesis for all clusters with a {\sc Sonority Angle} greater than $n$, with the result that either the cluster remains intact, or a competing solution to epenthesis emerges.

An example of {\sc Ident(Son$\measuredangle$)}$<$n blocking epenthesis is given in the sample tableaux below.

\ex. Ident(SonAngle)$<$1.5 $\gg$ {\sc *ComplexCoda} $\gg$ {\sc Dep}
     \a. Allows epenthesis in rising and level clusters such as stop-fricative TF:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/TF/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline a. & & TF & & $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & T\textipa{@}F & 0.59 & & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. Prevents epenthesis in falling clusters such as FT:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline c. & \ding{43} & FT & &\cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} \\
\hline d. & & F\textipa{@}T & 2.11 $\ast$! &\cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

I will illustrate the use of {\sc Ident(Son$\measuredangle$)}$<$n to block epenthesis in favour of other cluster resolution strategies with an example from the loanword phonology of the Dhaka dialect of Bangla, which is discussed and given an alternative analysis in \citep{karim.2011}. Standard Bangla keeps coda clusters in words borrowed from English, but these are simplified in the Dhaka dialect in a variety of ways, including both epenthesis and deletion of one of the consonants in the cluster.

@@

However, if we retain this ranking, we incorrectly expect RT clusters to be broken up by epenthesis rather than delete the [r]. This can be resolved by ranking {\sc Ident(Son$\measuredangle$)}$<$2.3, say, above {\sc Max}(r). This then blocks epenthesis in RT clusters, which have a {\sc Sonority Angle} of 2.35:

\ex. {\sc *ComplexCoda}, {\sc Ident(Son$\measuredangle$)}$<$2.3  $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{park}/} & {\sc *ComplexCoda} & {\sc Ident(Son$\measuredangle$)}$<$2.3 & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{park} & $\ast$! & & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline b. & & \textipa{parak} & & $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline c. & \ding{43} & \textipa{pak} & & & \cellcolor{lightgray}$\ast$ & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

While other analyses could possibly account for this data (see one such in \citep{karim.2011}), we see that simply ranking {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{F}$ allows us to account for it quite straightforwardly.

@@

Thus far, we have not made any particular use of free ranking. Although ranking \\ {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{M}$ does conflate all clusters with a {\sc Sonority Angle} above $n$, as well as all clusters with a {\sc Sonority Angle} below $n$, preserving only the distinction between {\sc Sonority Angles} greater and less than $n$, this could be accomplished using a universal ranking of the {\sc Ident(Son$\measuredangle$)}$<$n constraints as well, as illustrated in the following example:

\ex. The {\sc Ident(Son$\measuredangle$)}$<$n with the smallest $n$ that outranks {\sc *ComplexCoda} determines the cut-off point.

@@

Suppose a triconsonantal cluster /CCC/ has to be broken up either as [C\textipa{@}CC] or [CC\textipa{@}C], and further that [C\textipa{@}CC] violates some additional markedness constraint $\mathcal{M}$. Which {\sc Ident(Son$\measuredangle$)}$<$n constraint outranks $\mathcal{M}$ will then determine the positioning of the epenthetic vowel.

\ex. {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{M}$
     \a. Scenario 1: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
         Since neither violates {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n & $\mathcal{M}$ \\[0.5ex]
\hline \hline a. & \ding{43} & C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $<$n & \\
\hline b. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $<$n & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
      \b. Scenario 2: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$>$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
          C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} harmonically bound by C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} candidate.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline c. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}  & $<$n & \cellcolor{lightgray} \\
\hline d. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Scenario 3: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$<$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$>$n \\
         {\sc Ident(Son$\measuredangle$)}$<$n determines output, overruling $\mathcal{M}$.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline e. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$!  & \cellcolor{lightgray} \\
\hline f. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3}& $<$n  & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \d. Scenario 4: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2}-C\textsubscript{3}) $>$ n \\
         Since both candidates violate {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline g. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$  & \\
\hline h. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Because we have assumed that all other {\sc Ident(Son$\measuredangle$)}$<$n constraints rank low, certainly below $\mathcal{M}$, the degree to which candidates (g) and (h) violate {\sc Ident(Son$\measuredangle$)}$<$n is irrelevant --- the distinction between clusters with {\sc Sonority Angles} above $n$ has been conflated. Even if {\sc Sonority Angle}(C\textsubscript{2},C\textsubscript{3}) exceeds {\sc Sonority Angle}(C\textsubscript{1},C\textsubscript{2}), the outcome is determined by the markedness constraint $\mathcal{M}$.

@@

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc Ident(Son$\measuredangle$)}$<$1.5 &$ \mathcal{M}$  \\[0.5ex]
\hline \hline a. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} &  2.0 $\ast$!  & \cellcolor{lightgray} 2.0 $\ast$ & \cellcolor{lightgray} \\
\hline b.  & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} &  1.7  & \cellcolor{lightgray} 1.7 $\ast$ & \cellcolor{lightgray} $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Under a universal ranking of {\sc Ident(Son$\measuredangle$)}$<$n constraints, [C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}] would win in this scenario. In an equivalent scenario under free ranking, where {\sc Ident(Son$\measuredangle$)}$<$1.5 is the least stringent constraint to outrank $\mathcal{M}$, which in turn outranks all more stringent {\sc Ident(Son$\measuredangle$)}$<$n constraints, the [C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}] candidate would win instead. There would be conflation of all clusters with {\sc Sonority Angles} below $n$, but distinctions between clusters with {\sc Sonority Angles} would still be relevant.

The example of Chaha will show that free ranking is necessary, as conflation is observed on both sides of $n$, where {\sc Ident(Son$\measuredangle$)}$<$n is the most stringent constraint to outrank $\mathcal{M}$.

@@

As with {\sc Sonority Angle}, this gives rise to a family of constraints {\sc Ident(SonRise)}$<$n, defined similarly to {\sc Ident(Son$\measuredangle$)}$<$n.

@@

The fact that this process of epenthesis does not target obstruent-initial clusters is not problematic under the sonority contour faithfulness approach,
as the {\sc Ident(Son$\measuredangle$)} constraints limit epenthesis, rather than trigger it.  By choosing a more limited markedness constraint such as {\sc *Son-C},
defined below, we can have epenthesis act only on sonorant-initial clusters.

@@

Under a sonority contour faithfulness analysis, epenthesis is triggered by a markedness constraint on the structural conditions of the cluster, not by the markedness of the underlying sonority contour profile itself. Undominated {\sc *ComplexOnset} accounts for the non-existence of initial clusters. {\sc *Coda} is ranked low, and therefore medial clusters are left intact.{\sc *ComplexCoda}, dominated by an {\sc Ident(Son$\measuredangle$)} constraint, explains why some coda clusters are broken up by epenthesis while others are not.

@@

\ex. {\sc Ident(Son$\measuredangle$)}$<$1.5 $\gg$ {\sc *ComplexCoda}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{k1tf}/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} \\[0.5ex]
\hline \hline a. & & \textipa{k1tf} & & $\ast$! \\
\hline b. & \ding{43} & \textipa{k1t1f} & 0.59 & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{k1rm}/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} \\[0.5ex]
\hline \hline c. &  \ding{43} & \textipa{k1rm} & & $\ast$ \\
\hline d. & & \textipa{k1r1m} & 1.89 $\ast$! & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc Ident(Son$\measuredangle$)}$<$1.9 $\gg$ {\sc *ComplexCoda}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{k1rm}/} & {\sc Ident(Son$\measuredangle$)}$<$1.9 & {\sc *ComplexCoda} \\[0.5ex]
\hline \hline a. &   & \textipa{k1rm} & & $\ast$! \\
\hline b. & \ding{43} & \textipa{k1r1m} & 1.89 & \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
